---
title: "p8108_hw3_yz4436"
author: "Yuanhao Zhang"
date: "2022-10-13"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(ggridges)
library(patchwork)

library(p8105.datasets)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

### Problem 1

#### Read in the data

```{r}
data("instacart")

instacart = 
  instacart %>% 
  as_tibble(instacart)
```

#### Answer questions about the data

This dataset contains `r nrow(instacart)` rows and `r ncol(instacart)` columns, with each row resprenting a single product from an instacart order. Variables include identifiers for user, order, and product; the order in which each product was added to the cart. There are several order-level variables, describing the day and time of the order, and number of days since prior order. Then there are several item-specific variables, describing the product name (e.g. Yogurt, Avocado), department (e.g. dairy and eggs, produce), and aisle (e.g. yogurt, fresh fruits), and whether the item has been ordered by this user in the past. In total, there are `r instacart %>% select(product_id) %>% distinct %>% count` products found in `r instacart %>% select(user_id, order_id) %>% distinct %>% count` orders from `r instacart %>% select(user_id) %>% distinct %>% count` distinct users.


Below is a table summarizing the number of items ordered from aisle. In total, there are 134 aisles, with fresh vegetables and fresh fruits holding the most items ordered by far.

```{r}
instacart %>% 
  count(aisle) %>% 
  arrange(desc(n))
```

Next is a plot that shows the number of items ordered in each aisle. Here, aisles are ordered by ascending number of items.

```{r}
instacart %>% 
  count(aisle) %>% 
  filter(n > 10000) %>% 
  mutate(aisle = fct_reorder(aisle, n)) %>% 
  ggplot(aes(x = aisle, y = n)) + 
  geom_point() + 
  labs(title = "Number of items ordered in each aisle") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))

```

Our next table shows the three most popular items in aisles `baking ingredients`, `dog food care`, and `packaged vegetables fruits`, and includes the number of times each item is ordered in your table.

```{r}
instacart %>% 
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>%
  group_by(aisle) %>% 
  count(product_name) %>% 
  mutate(rank = min_rank(desc(n))) %>% 
  filter(rank < 4) %>% 
  arrange(desc(n)) %>%
  knitr::kable()
```

Finally is a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week. This table has been formatted in an untidy manner for human readers. Pink Lady Apples are generally purchased slightly earlier in the day than Coffee Ice Cream, with the exception of day 5.

```{r}
instacart %>%
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>%
  group_by(product_name, order_dow) %>%
  summarize(mean_hour = mean(order_hour_of_day)) %>%
  spread(key = order_dow, value = mean_hour) %>%
  knitr::kable(digits = 2)
```


### Problem 2

Import and tidy the dateset accel_data.csv. I create a variable weekday_vs_weekend to indicate the day is weekday(Mon-Fri) or weekend(Sat-Sun), then I relocate weekday_vs_weekend after variable day. Finally, I use pivot_longer to write from activity_1 to activity_1440 into variable Activity_list, and values to Activity_datavalues.
```{r}
accel_data_df =
  read_csv(
    "data/accel_data.csv",
    show_col_types = FALSE) %>%
  janitor::clean_names() %>%
  mutate(
    weekday_vs_weekend = day,
    weekday_vs_weekend = recode(weekday_vs_weekend, "Monday" = "Weekday", 
                                "Tuesday" = "Weekday", "Wednesday" = "Weekday", 
                                "Thursday" = "Weekday", "Friday" = "Weekday", 
                                "Saturday" = "Weekend", "Sunday" = "Weekend")
  ) %>%
  relocate(week, day_id, day, weekday_vs_weekend) %>%
  pivot_longer(
    activity_1:activity_1440,
    names_to = "Activity_list", 
    values_to = "Activity_datavalues")
```

* Description about the resulting dataset:
  * The number of variables in the resulting dateset accel_data_df is `r ncol(accel_data_df)`, which are `r colnames(accel_data_df)`.
* The number of observations in the resulting dataset combine_tidy is `r nrow(accel_data_df)`.


I aggregate across minutes to create a total_activity variable for each day, and create a table showing these totals.
```{r}
accel_data_df %>%
  group_by(week, day_id, day) %>%
  summarize(total_activity = sum(Activity_datavalues)) %>%
  knitr::kable()
```
* I do not see any trends apparent based on the table that showing totals.


Make a single-panel plot that shows the 24-hour activity time courses for each day and use color to indicate day of the week.
```{r}
accel_data_df %>%
  separate(Activity_list, into = c("Activity", "minute"), sep = "_") %>%
  mutate(
    minute = as.numeric(minute)
  ) %>%
  ggplot(aes(x = minute, y = Activity_datavalues, color = day)) + 
  geom_point(alpha = .5) + labs(y = "activity counts", title = "24-hour activity time courses for each day") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1)) + 
  theme(axis.text.y = element_text(angle = 60, hjust = 1)) + 
  scale_x_continuous(
    breaks = seq(0, 1440, 60), 
    limits = c(0, 1440))
```
* Based on this graph, we could find that there is no specific pattern. Also, the highest value appears around 1140 to 1200 minutes, which is 7pm to 8pm, and the lowest value appears around 0 to 240 minutes, which is 0am to 4am.


### Problem 3

* The size of the dataset is that there are `r nrow(ny_noaa)` rows and `r ncol(ny_noaa)` columns, which are `r colnames(ny_noaa)`. There are a lot of NAs and data missing is a big issue for this dataset.


Import and tidy the dateset ny_noaa. Then I create separate variables for year, month, and day, and make reasonable units or temperature, precipitation, and snowfall. 
```{r}
data("ny_noaa")
ny_noaa =
  ny_noaa %>%
  janitor::clean_names() %>%
  #drop_na() %>%
  separate(date, into = c("year", "month", "day"), sep = "-") %>%
  mutate(
    prcp = prcp/10,
    tmax = as.numeric(tmax)/10,
    tmin = as.numeric(tmin)/10,
    snow = snow/100
  )
ny_noaa %>%
  count(snow, name = "observed values")
```
* For snowfall, the most commonly observed values is 0 since I find the mode of snow is 0.

Make a two-panel plot showing the average max temperature in January and in July in each station across years. 
```{r}
ny_noaa %>%
  filter(month %in% c("01", "07")) %>%
  group_by(id, month, year) %>%
  summarize(avg_tmax = mean(tmax, na.rm = TRUE))  %>%
  ggplot(aes(x = year, y = avg_tmax)) + 
  geom_point(aes(color = id), size = 0.1, alpha = 0.5) +
  labs(title = "The average max temperature in January and in July in each station across years") + facet_grid(. ~ month) + 
  theme(legend.position = "none") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1)) + 
  theme(axis.text.y = element_text(angle = 60, hjust = 1))
```
* Through the plot, we could see that the for January, the average max temperature is from -10 to 10 and that for July is from 20 to 35. We could also find that the average max temperature in January spread wider than the average max temperature in July. According to the plot, we could see that they all have outliers, since for both January and July, there are some points that far from the mean part.
